<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{A guide for importing and analysing ruminal pH bolus data}
-->

```{r, echo = FALSE, include=FALSE}
set.seed(2018-08-06)
```

---

<center> <h1>A guide for importing and analysing ruminal pH bolus data</h1> </center> 

---

<center> <h2>Introduction</h2> </center> 

---

TODO


---

<center> <h2>Installation</h2> </center> 

---

Install the required packages from CRAN, as well as the BoluspH data from GitHub (this only needs to be done once, unless the package is updated):

```{r, eval=FALSE}
install.packages(c('tidyverse','mgcv','devtools'))
devtools::install_github('boydorr/BoluspH')
```

You should then be able to access this vignette from within R using:

```{r, eval=FALSE}
vignette('BoluspH', package='BoluspH')
```


---

<center> <h2>Basic Example</h2> </center> 

---

Let's say we have 3 files to import - one SmaXtec file, one WellCow file and one in a custom format.  First we need to set up a data frame with (at a minimum) the ID numbers to be used plus the filenames relating to these individuals:

```{r}
meta <- data.frame(ID=c('Cow_1','Cow_2','Cow_3'), Filename=c('ST1.xlsx','WC1.csv','CS1.csv'), stringsAsFactors=FALSE)
```

This could optionally also include milking times, start and end dates for the data to be used, and other information e.g. Farm that will be merged with the outputs, e.g.:

```{r}
meta <- data.frame(ID=c('Cow_1','Cow_2','Cow_3'), Filename=c('ST1.xlsx','WC1.csv','CS1.csv'), Milking=c('09:00,16:00', '07:00,13:00,18:00', ''), Farm=c('FarmA','FarmB','FarmC'), stringsAsFactors=FALSE)
```

We then need to set up a model container using this meta-data:

```{r}
library('BoluspH')
model <- BoluspH(meta)
```

Then we can add the data by telling the model the working directory in which it can be found (in this case within the package itself for demonstration purposes).  First we can add SmaXtec data (these are the only ones with .xlsx extension), and then WellCow data for individual 'Cow_2', and then finally the rest of the data in a custom data format:

```{r}
testdir <- system.file("extdata", package="BoluspH")
model$AddSmaXtecData(testdir)
model$AddWellCowData(testdir, ID='Cow_2')
model$AddCustomData(testdir, skip=13, date_col=3, time_col=4, pH_col=6, csv=TRUE, sep=';', dec=',', date_format='%d.%m.%Y', time_format='%H:%M')
```

It is also possible to add data from different paths using multiple calls to the methods above, and/or add data in different formats using the AddCustomData method.

Once all data has been imported we run the models, optionally including the GAM (long-term trend) and milking-related sine wave terms ('none', 'fixed' or 'adjusted'):

```{r}
model$RunModels(gam=TRUE, milking='adjusted')
```

Notice that Cow_1 was automatically skipped due to insufficient data (this can be changed by adjusting the exclude_pre, exclude_post, and min_days arguments to the RunModels method).

Then extract and look at the overview information:

```{r}
overview_out <- model$GetOverview()
overview_out
```

And then extract the full observation-level data including predictions and residuals:

```{r}
data_out <- model$ExtractData()
str(data_out)
summary(data_out)
head(data_out)
```

Then further analyses can be done e.g.:

```{r}
ggplot(data_out, aes(x=Time)) +
	geom_line(mapping=aes(y=pH)) +
	geom_line(mapping=aes(y=prediction), col='blue', lwd=2) +
	facet_wrap(~ID, scale='free')
```

Or daily mean absolute residuals can be calculated e.g.:

```{r}
daily <- data_out %>% group_by(ID, Date) %>%
	summarise(mean_abs_resid = mean(abs(residual)))
head(daily)

ggplot(daily, aes(x=Date, y=mean_abs_resid)) +
	geom_line() +
	facet_wrap(~ID, scales='free_x')
```

The daily data for a specific animal can also be saved to CSV e.g.:

```{r}
data_ind <- model$ExtractData(ID='Cow_2')
str(data_ind)
summary(data_ind)

write.csv(data_ind, file='Cow_2.csv', row.names=FALSE)
```


---

### This vignette was built with:

```{r}
sessionInfo()
```
